---
title: "Assignment 2"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "")
```

```{r, include = F}
#### if need to install packages, run below
# install.packages("here")
# install.packages("readr")
# install.packages("tidyverse")
# install.packages("table1")
# install.packages("ggplot2")
```

#### Imports
```{r, message = F, warning = F}
library(here)
library(tidyverse)
library(readr)
library(table1)
library(ggplot2)
```

#### Load Data
```{r}
raw_data <- read_csv(here("raw-data", "cohort.csv"))
glimpse(raw_data)
```

#### 1. Table
```{r}
df <- 
    raw_data |> 
    mutate(
        smoke = factor(smoke, 
                       levels = c(0,1), 
                       labels = c("no", "yes")), 
        female = factor(female, 
                        levels = c(0,1), 
                        labels = c("no", "yes")), 
        cardiac = factor(cardiac, 
                         levels = c(1,0), 
                         labels = c("cardiac event", "no cardiac event"))
    ) |> 
    mutate(
        age_cat = case_when(
        age < 30 ~ "18-29", 
        age >= 30 & age < 50 ~ "30-49", 
        age >= 50 ~ "50+", 
        TRUE ~ NA_character_
    ))


tb1 <- table1(~ age_cat + female + cost + smoke + cardiac, data = df)
tb1
```

#### 2. Analysis: linear regression
```{r}
model <- glm(cost ~ age_cat + female + smoke + cardiac, data = df)
summary(model)
```

#### 3. Figure: 
```{r}
df |> 
    ggplot() + 
    geom_boxplot(aes(x = age_cat, y = cost, color = age_cat, fill = age_cat), 
                 alpha = 0.2) + 
    labs(x = "Age Category (years)", 
         y = "Cost (dollars)", 
         color = NULL, fill = NULL) 
    
```


#### 4. Summary:
I ran a linear regression to assess the relationship of `age` and `cost`, adjusting for `female`, `smoke`, and `cardiac`. `age` was categorized into three groups: 18-29, 30-49, and 50+ for better visualization and interpretability. From our linear model, we see that age is a significant and positive predictor of `cost` and that individuals in older age groups are more likely to pay more `cost` than younger age groups (on average, there seems to be ~200-300 increase in `cost` per age group). The figure shows the distribution of `cost` by `age_cat`.

### AI Statement:
I did not use any generative AI technology to complete any portion of the work.
